---
title: "Dorchester and Boston Crime Data"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Scrape Data from online
2. Exploratory Analysis


About the data - Universal Hub, started in 2005, is a community news and information site for the Boston area based on the idea that residents can help keep themselves informed. Combining reports from a pool of hundreds of local bloggers, photographers and tweeters, Universal Hub posts daily news and links, often before the mainstream media, and gives local folks a place to discuss the day's events - and to post news on their own.
```{r include=FALSE}
# Load packages
library(tidyverse)
require(readr)
require(dplyr)
require(ROCR)
require(rvest)
```

```{r include=FALSE}
# Scrape the neighbourhood nodes
response <- read_html("https://www.universalhub.com/crime/home.html")
neighborhood <- rvest::html_nodes(x = response,
                                  xpath = '//div[@class="panel-pane pane-block pane-jump-menu-jump-menu-m-menu-neighborhood-cr"]//option[@class="d-1"]')

## Get the neighbourhood names
nb_names <- rvest::html_text(neighborhood)

## Get the url of each neighbourhood
nb_url <- rvest::html_attr(neighborhood, name = "value")
nb_url <- paste0("https://www.universalhub.com", nb_url)


# Write a function to scrape the variables from the url
scrape_crime <- function(url) {
  response <- read_html(url)
  
  # Type of the crime
  crime <- rvest::html_nodes(x = response, 
                            xpath = '//td[contains(@class, "name")]') %>%
    rvest::html_text(trim = T)
  # Time of the crime
  hour <- rvest::html_nodes(x = response, 
                            xpath = '//span[contains(@class, "date")]') %>%
    rvest::html_text() %>%
    lubridate::parse_date_time(orders = "%m/%d/%y - I:M p") %>%
    lubridate::hour()
  
  return(data.frame(crime, hour))
}


# Write a function to extract the url of the next page
get_next_page <- function(url) {
  initial_page <- read_html(url)
  # Extract the url of the next page
  next_page <- rvest::html_nodes(x = initial_page,
                                 xpath = '//ul[@class="pager"]//a[@title = "Go to next page"]') %>%
    html_attr(name = "href")
  # If there is a next page (the url exists), then return this url
  # If there is no next page (the url does not exist, then return NA)
  if (length(next_page) > 0) {
    next_page <- paste0("https://www.universalhub.com", next_page)
    return(next_page) 
  } else {
    return(NA)
  }
}


# Write a function to read each page of the neighbourhood
scrape_each_page <- function(page1_url) {
  
  # Scrape the data from page 1
  crime_data <- scrape_crime(page1_url)
  
  # Extract the url of the next page
  nextpage_url <- get_next_page(page1_url)
  
  # While there is a next page, go to the next page and scrape the data
  while (!is.na(nextpage_url)) {
    ## Scrape the data
    nextpage_data <- scrape_crime(nextpage_url)
    ## Combine the data to the data at the previous page
    crime_data <- rbind(crime_data, nextpage_data)
    ## Get the url of another next page
    nextpage_url <- get_next_page(nextpage_url)
  }
  
  return(crime_data)
}


# Create a tibble to store the results for all the neighbourhoods
crime_data <- scrape_each_page(nb_url[1]) %>%
  mutate(neighborhood = nb_names[1])
for (i in 2:length(nb_url)) {
  crime_temp <- scrape_each_page(nb_url[i]) %>%
    mutate(neighborhood = nb_names[i])
  crime_data <- rbind(crime_data, crime_temp)
}


# Save the data
#write_csv(crime_data, "dorchester_crime_data.csv")

```

## Clean Data


```{r}
# Clean crime 
crime_data <- crime_data %>% mutate(crime = tolower(crime)) # make crimes all lowercase
 
shooting_vec <- c("shooting", "shootin", "shoting", "shotting") #all the shooting spellings 
crime_data$crime <- ifelse(crime_data$crime %in% shooting_vec, "shooting", crime_data$crime) 

```


The Top 5 Most Common Crimes in Dorchester, MA
```{r}
crime_count <- crime_data %>% 
  group_by(crime) %>% 
  summarize (count = n()) %>% 
  arrange(desc(count))
top_5_crimes = crime_count %>% slice(1:5)
top_5_crimes

```

Number of Reported Crimes by Hour
```{r}
crimes_by_hour <- crime_data %>% 
  group_by(hour) %>% 
  summarize (count = n())

ggplot(crimes_by_hour, aes(x=hour, y=count)) +
  geom_line(color = "blue", linetype="solid", size =2) + theme_classic() + 
  labs(title="Total number of reported crimes by hour", x="Hour of the Day", 
       y = "Total number of reported crimes")
```

The total number of reported crimes by hour is interestingly U-shaped. There is a drop in the number of crimes reported in the early morning and then steadily increases throughout the day with the most amount of crimes occurring during the hours of 9pm to 11 pm. 

```{r}
top_5 <- top_5_crimes$crime
crimes_by_hour_type <- crime_data %>% 
  group_by(hour, crime) %>% 
  summarize (count = n()) %>% 
  filter(crime %in% top_5)

ggplot(crimes_by_hour_type, aes(x=hour, y=count, color = crime)) +
  geom_line(linetype="solid", size = 1) + theme_classic() + 
  labs(title="Total number of reported crimes by hour for top 5 crimes in Boston", 
       x="Hour of the Day", y = "Total number of reported crimes")
```
Although the five crimes types are nested together in the daytime, the volume of reported shootings increases after late afternoon and becomes the highest in the evening.

```{r}
crime_by_hour_neighborhood <- crime_data %>% 
  group_by(hour, neighborhood) %>% 
  summarize (count = n()) %>% 
  filter(neighborhood %in% c("Downtown", "Dorchester"))

ggplot(crime_by_hour_neighborhood , 
                                  aes(x=hour, y=count, color = neighborhood)) +
  geom_line(linetype="solid", size = 1) + theme_classic() + 
  labs(title="Reported Crimes by hour in Dorchester and Downtown Boston", 
       x="Hour of the Day", y = "Total number of reported crimes")
```
There is less reported crime in Downtown Boston than in Dorcester. In the early morning there is a decrease in reported crime in Dorcester and it slowly increase throughout the day peaking at night. In Downtown Boston, crime is reported  most in the late afternoon and then drops in evening.





